---
title: 'FP1：public transportation'
author: "Jenny Li"
output: 
  html_document:
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

## Research question
- Where are there gaps in a public transportation system?
- What are the connections between public transport systems and    environmental protection and emissions? 
- How much money the government spends (or future spends) on transportation based energy cost ?

## Dataset
```{r libraries}
library(tidyverse)     # for graphing and data cleaning
library(lubridate)     # for date manipulation
library(ggthemes)      # for even more plotting themes
library(dplyr)
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
Sys.setlocale("LC_TIME", "English")
theme_set(theme_minimal())
```

### Data 1
```{r data1}
calendar<-read.delim('calendar_dates.txt', sep=',')
stop_times<-read.delim('stop_times.txt', sep=',')
stops<-read.delim('stops.txt', sep=',')
trips<-read.delim('trips.txt', sep=',')
routes<-read.delim('routes.txt', sep=',')
```

#### Clean Data
```{r}
newTrips <- trips %>%
  select(-trip_long_name:-bikes_allowed)
newStop <- stops %>% 
  select(-location_type:-zone_id)
newRoutes<-routes %>% 
  select(c(route_id,route_long_name))
newStopTimes<- stop_times %>% 
  select(-stop_headsign) %>% 
  select(-pickup_type:-fare_units_traveled) 
newCalendar <- calendar %>% 
  select(-exception_type)
```

#### Combine Data
```{r}
walloonTransit<-newStopTimes %>% 
  left_join(newStop,by="stop_id") %>% 
  left_join(newTrips, by="trip_id") %>% 
  left_join(newCalendar, by="service_id") %>% 
  left_join(newRoutes, by="route_id") %>% 
  filter(date > 20180731& arrival_time<"06:00:00")
```

#### Data Source
[TEC](http://gtfs.ovapi.nl/tec/gtfs-tec-latest.zip)
#### Data Description
Transit data such as stop locations, scheduled times and dates, etc.
#### Data Limitation
The data was collected 4 years ago. Thus is a little outdated and also the data does not contain the measure passenger flow. Also, since the dataset is too large, I decided to limit it to the early hours of each day in August.
#### Variables
```{r}
head(walloonTransit)
```
stop_lon and stop_lat including the arrival_time of the trips are interesting. The stop_lon and stop_lat are basically the lontitude and latitude of each stop and arrival_time is the time when the train arrivals at the stop.
#### Visualization
```{r, fig.alt="map for the stops"}
mapStops <- get_stamenmap(
  bbox = c(left = 2, bottom = 49.4, right = 6, top = 51), 
  maptype = "toner",
  zoom = 12)

ggmap(mapStops) +
  geom_point(
    data = walloonTransit,
    aes(x = stop_lon, y = stop_lat),
    alpha = .3,
    size = .1,
    color = "maroon4"
  ) +
  labs(title = "Stops In Wallon Belgium")
```
#### Join Variables
All the variables including "id" would be possible to join with other dataset, such as trip_id or route_id.

