---
title: 'FP1'
author: "Jenny Li"
output: 
  html_document:
    keep_md: TRUE
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```
## Project Area
Public Transportation

## Members
Jenny, Pippa, Rita


## Research question
- Where are there gaps in a public transportation system?
- What are the connections between public transport systems and    environmental protection and emissions? 
- How much money the government spends (or future spends) on transportation based on energy cost ?

## Dataset
```{r libraries}
library(tidyverse)     # for graphing and data cleaning
library(lubridate)     # for date manipulation
library(ggthemes)      # for even more plotting themes
library(dplyr)
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(readr)
library(readxl)
library(imputeTS)
library(plotly)
Sys.setlocale("LC_TIME", "English")
theme_set(theme_minimal())
```

### Data 1
```{r data1}
calendar<-read.delim('calendar_dates.txt', sep=',')
stop_times<-read.delim('stop_times.txt', sep=',')
stops<-read.delim('stops.txt', sep=',')
trips<-read.delim('trips.txt', sep=',')
routes<-read.delim('routes.txt', sep=',')
```

#### Clean Data
```{r}
newTrips <- trips %>%
  select(-trip_long_name:-bikes_allowed)
newStop <- stops %>% 
  select(-location_type:-zone_id)
newRoutes<-routes %>% 
  select(c(route_id,route_long_name))
newStopTimes<- stop_times %>% 
  select(-stop_headsign) %>% 
  select(-pickup_type:-fare_units_traveled) 
newCalendar <- calendar %>% 
  select(-exception_type)
```

#### Combine Data
```{r}
walloonTransit<-newStopTimes %>% 
  left_join(newStop,by="stop_id") %>% 
  left_join(newTrips, by="trip_id") %>% 
  left_join(newCalendar, by="service_id") %>% 
  left_join(newRoutes, by="route_id") %>% 
  filter(date > 20180731& arrival_time<"06:00:00")
```

#### Data Source
[TEC](http://gtfs.ovapi.nl/tec/gtfs-tec-latest.zip)

#### Data Description
Transit data such as stop locations, scheduled times and dates, etc.

#### Data Limitation
The data was collected 4 years ago. Thus is a little outdated and the data does not contain the measure passenger flow. Also, since the dataset is too large, I decided to limit it to the early hours of each day in August.

#### Variables
```{r}
head(walloonTransit)
```

stop_lon and stop_lat including the arrival_time of the trips are interesting. The stop_lon and stop_lat are basically the lontitude and latitude of each stop and arrival_time is the time when the train arrivals at the stop.

#### Visualization
```{r, fig.alt="map for the stops"}
mapStops <- get_stamenmap(
  bbox = c(left = 2, bottom = 49.4, right = 6, top = 51), 
  maptype = "toner",
  zoom = 12)

ggmap(mapStops) +
  geom_point(
    data = walloonTransit,
    aes(x = stop_lon, y = stop_lat),
    alpha = .3,
    size = .1,
    color = "maroon4"
  ) +
  labs(title = "Stops In Wallon Belgium")
```

#### Join Variables
All the variables including "id" would be possible to join with other dataset, such as trip_id or route_id.

### Data 2

```{r}
energyUse<- read_excel("Energy Use.xls", sheet="Data") %>% 
  na_replace(0) %>% pivot_longer(
                cols = c("1960":"2021"),
                names_to = "year",
                values_to = "country_energy_used"
)
```

#### Data Source
[OECD](https://data.worldbank.org/indicator/EG.USE.PCAP.KG.OE)

#### Data Description
Energy use (kg of oil equivalent per capita) by countries since 1960 through 2015.

#### Data Limitation
The data was collected 6 years ago. Thus is a little outdated and the data only contains one kind of energy that is used by countries. In addition, some of the data is missing.

#### Variables
```{r}
head(energyUse)
```

Country Name, year and country_energy_used are the three interesting variables in this dataset. Country Name refers to the name of the country, year refers to year and country_energy_used refers to the use of energy (kg of oil equivalent per capita) by each country.

#### Visualizaiton
```{r, fig.alt="Energy Comsuption By Countries"}
energyUseGraph<- energyUse %>%
  ggplot(aes(x = year, y = country_energy_used)) +
  labs(title = "Energy Comsuption By Countries", y = "Comsuption (kg of oil equivalent per capita)", x =
         NULL) +
  geom_point(aes(color = `Country Name`)) +
  theme(
    plot.title = element_text (hjust = 0.5,
                               face = "bold",
                               size = 11),
    axis.text.x = element_text(
      size = 6,
      angle = -90,
      hjust = 0
    ),
    panel.grid.major.x  = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.position = 'none'
  )

ggplotly(energyUseGraph, tooltip = c("Country Name","colour") )
```

#### Join Variables
Country Name, Country Code and year are possible to join with other dataset.